### 2020-12-16 Thesis-Chapter2-Background.tex
Address some TODOs

A short law designed to grant federal courts the right to ``issue all writs necessary or appropriate in aid of their
respective jurisdictions and agreeable to the usages and principles of law,'' creating a legal framework for providing
data to law enforcement. The \acs{AWA} was famously cited in the Apple vs. FBI case, though its applicability to \ac{EA}
in general is unsettled \cite{matyas_incommensurability_2018}.

@@

The most frequently cited constitutional amendments in the \ac{EA} debate. I: freedom of speech; IV: protection from
unreasonable searches and seizures and requirements for warrants; V: right against self-incrimination.

@@

The \ac{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
Provides explicit protections for private electronic communications data, and updated definitions to contemporary
technology. It grants law enforcement explicit access to certain data as well; some terminology has aged poorly with
technological evolution, making the access increasingly permissive \cite{shamsi_2011}.

}
% FIXME: make tense for enacted laws present

@@

\ac{CALEA} mandates that telecommunications carriers enable the government to ``intercept all of the subscriber's wire
and electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized
to be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant
that the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
specifically excludes ``information services'' from its interception requirement which has included internet platforms,
though that exclusion is currently under attack \cite{pfefferkorn_2019}.

@@

% TODO: find a "con" word that means "denied"! Contested?

The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
authoritative legal texts. However, whether it is condemned, denied, conceded, or condoned, the judicial branch also
establishes legal doctrine and actively enforces rulings in a manner that amounts to creation of public policy
\cite{feeley_judicial_2000}. Due to Congress's clear authority over telecommunications and federal investigatory powers,
the courts are most likely to exert influence in this area through interpretation of current law. This interpretation
will both determine how present law is enforced and draw constitutional boundaries around any actions the executive or
legislative branches undertake.

>>>>

The \acs{AWA} is a short law designed to grant federal courts the right to ``issue all writs necessary or appropriate in
aid of their respective jurisdictions and agreeable to the usages and principles of law,'' creating a legal framework
for providing data to law enforcement. The \acs{AWA} was famously cited in the Apple vs. FBI case, though its
applicability to \ac{EA} in general is unsettled \cite{matyas_incommensurability_2018}.

@@

Three Bill of Rights amendments are frequently cited in the \ac{EA} debate. I: freedom of speech; IV: protection from
unreasonable searches and seizures and requirements for warrants; and V: the right against self-incrimination.

@@

\acs{ECPA} includes the Stored Communications Act and Pen Register Act, amending the switchboard-era Wiretap Act.
Provides explicit protections for private electronic communications data, and updated definitions to contemporary
technology. It grants law enforcement explicit access to certain data as well; some terminology has aged poorly with
technological evolution, making the access increasingly permissive \cite{shamsi_2011}.

@@

\acs{CALEA} mandates that telecommunications carriers enable the government to ``intercept all of the subscriber's wire
and electronic communications'' in a manner that ``protects the privacy and security'' of communications not authorized
to be intercepted. The law does not authorize law enforcement to compel use of any specific technologies, which meant
that the \ac{clipperchip} proposal issued a year earlier could not be made mandatory through \acs{CALEA}. \acs{CALEA}
specifically excludes ``information services'' from its interception requirement which has included internet platforms,
though that exclusion is currently under attack \cite{pfefferkorn_2019}.

@@

The primary judiciary function is to adjudicate the law through the determination of facts and interpretation of
authoritative legal texts. However, whether it is contested, conceded, condemned, or condoned, the judicial branch also
establishes legal doctrine and actively enforces rulings in a manner that amounts to creation of public policy
\cite{feeley_judicial_2000}. Due to Congress's clear authority over telecommunications and federal investigatory powers,
the courts are most likely to exert influence in this area through interpretation of current law. This interpretation
will both determine how present law is enforced and draw constitutional boundaries around any actions the executive or
legislative branches undertake.

### 2020-12-16 Thesis-Chapter4-Tools.tex
Address some TODOs

% TODO: can probably cut more
Argument mapping tools originated in 1970 with text-based \acp{IBIS} \cite{kunz_issues_1970}. Interest in \acp{IBIS} has
waxed and waned. Some software tools have been developed, including the now-defunct gIBIS \cite{conklin_gibis_1988} and
Compendium \cite{dutoit_hypermedia_2006}, which were designed to produce graphical argument representations. Argdown is
a recent tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
\cite{voigt_argdown_2018}. Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA}
debate.

>>>>

Argument mapping tools originated in 1970 with text-based \acp{IBIS} \cite{kunz_issues_1970}. Graphical software mapping
tools include the now-defunct gIBIS \cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006}. Argdown is a
recent tool that generates graphical \acp{argmap} from structures specified in a Markdown-based language
\cite{voigt_argdown_2018}. Argdown is used in \mychap{chap-arguments} to analyze the arguments used in the \ac{EA}
debate.

### 2020-12-16 Thesis-Chapter5-Arguments.tex
Address some TODOs

\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, \ac{encryption}'s dual
contributions to information security and radical privacy are central to the debate. Of course, there is only conflict
when radical privacy is perceived as a negative. The right to privacy is both a strongly held value and an enshrined
legal principle. Many use privacy concerns to argue that \ac{EA} is socially undesirable. The Snowden revelations
\cite{landau_making_2013} unveiled the scale of privacy-eroding U.S. government surveillance enabled by technological
changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract worry
to many, but for the most vulnerable, they are frighteningly concrete. For example, Saudi journalist Jamal Khashoggi's
fell
% TODO: bookmark

private communications were surveilled by the Saudi government before he became a victim of politically motivated murder
\cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free speech
\cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.

@@

% TODO: I still really want to use this!
% European law enforcement's infiltration of encrypted messaging service Encrochat exposed crime rings and led to a
% massive raids \cite{cox_2020}.

@@

\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA} \cite{bellovin_lawful_2013}
\cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018} \cite{kerr_encryption_2017} \cite{soesanto_2018}.
\Ac{lawful-hacking}'s support may be due in part to the fact that it is a technical approach that isn't \ac{EA} itself.
However, its effectiveness against criminals for whom \ac{EA} would not work is its obvious benefit. This is because,
while organized crime and terror groups would still use encryption tools to evade an \ac{EA} scheme, their operational
security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
This approach is already in use \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
\cite{feuer_chapo_2019}.

The real change that \ac{lawful-hacking} would take would be a sanctioning and formalization of this strategy.
% TODO: tie to paragraph above?

>>>>

\myfig{fig-args-contrib} maps the central arguments in the EA debate. As previously stated, \ac{encryption}'s dual
contributions to information security and radical privacy are central to the debate. Of course, there is only conflict
when radical privacy is perceived as a negative. The right to privacy is both a strongly held value and an enshrined
legal principle. Many use privacy concerns to argue that \ac{EA} is socially undesirable. The Snowden revelations
\cite{landau_making_2013} unveiled the scale of privacy-eroding U.S. government surveillance enabled by technological
changes, terrorism-motivated policies, and weak oversight \cite{shamsi_2011}. Privacy violations are an abstract worry
to many, but for the most vulnerable, they are frighteningly concrete. For example, mobile phone surveillance malware
enabled the Saudi government to capture and murder journalist Jamal Khashoggi for expressing dissent against the crown
prince \cite{liebermann_2019}. At a societal level, the mere presence of surveillance changes behavior and chills free
speech \cite{rogaway_moral_2015}, and government violations of the law degrade institutional trust.

@@

% TODO: Possibly find a new spot for this Encrochat story
% European law enforcement's infiltration of encrypted messaging service Encrochat exposed crime rings and led to a
% massive raids \cite{cox_2020}.

@@

\Ac{lawful-hacking} is the most prominent proposed alternative to \ac{EA} \cite{bellovin_lawful_2013}
\cite{hennessey_lawful_2016} \cite{rozenshtein_wicked_2018} \cite{kerr_encryption_2017} \cite{soesanto_2018}.
\Ac{lawful-hacking}'s support may be due in part to the fact that it is a technical approach that isn't \ac{EA} itself.
However, its effectiveness against criminals for whom \ac{EA} would not work is its obvious benefit. This is because,
while organized crime and terror groups would still use encryption tools to evade an \ac{EA} scheme, their operational
security would fail at some point, letting law enforcement or intelligence agencies into their networks and devices.
This approach is already in use \cite{cox_2020}, most famously to bring down cartel kingpin El Chapo
\cite{feuer_chapo_2019}. Because it is already used, deliberately choosing \ac{lawful-hacking} as the primary strategy
for accessing plaintext would mean officially sanctioning, formalizing, and funding \ac{lawful-hacking} programs.
